Indirect Conflicts: An Exploration and Discussion of Tools, Process, and Developer Insight

### 2013-11-06 iced-2013/paper.tex
ICED new introduction. It is limited to a single page now.

As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution, however,
in practice we see these mitigated losses continue to appear quite frequently. Not only do these conflicts still occur,
but at times they can prove to be a significant and time-consuming chore for developers~\cite{Perry:2001:PCL}.

Two types of conflicts have attracted the attention of academics over the years, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact, or finding expert knowledge about a particular file. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. However, indirect conflict tools have not shared the same success. Indirect conflicts
can arise in source code for such reasons as having one's own code negatively affected by a library upgrade,
making a code change only to find out it had negative effects on unit or integration test, or having negative effects
on the way other developer's have interpreted the use of your code. Tools that have attempted to either
help developers prevent or solve indirect conflicts have either attracted little developer interest or have
had some common struggles which remain to be studied or solved.

Difficulties regarding indirect conflicts are to determine when software dependencies change and more crucially,
when those changes will create negative consequences in a software project.
Sarma et al.~\cite{Sarma:2007:TSA} built Palantir, which can both detect potential indirect conflicts, at the class signature level,
and alert developers to these conflicts. Holmes et al.~\cite{Holmes:2010:CAR} take it
one step further with their tool YooHoo by detecting fine grained source code changes, such as method return type changes, 
and create a taxonomy for different
types of changes and their proneness to cause indirect conflicts. The tool Ariadne~\cite{Trainer:2005:BGT} 
creates an environment where developers can see how source code changes will affect other areas of a project at the 
method level and thus where indirect conflicts may occur. 
Another indirect conflict tool, CASI~\cite{Servant:2010:CPI},
utilizes dependency slicing~\cite{Bajracharya:2009:SIS} instead of method call graphs to provide 
an environment to see what areas of a project are being affected by a source code change. 
Most of these tools have all shown to have common difficulties of: scalability, false positives, information overload, 
or providing redundant information to preexisting tools. 
Our own tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered these same fates.

Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
However, socio-technical congruence is largely unproven in regards to its correlation to
software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.

We are interested in exploring three questions in the hopes of resolving some of the issues that are occurring
with indirect conflict tools:

@@

We interviewed 19 developers from across 12 companies and open source projects as well as surveyed 78 
developers in order to answer the aforementioned questions. Our findings indicate that while indirect conflicts are very much a real 
and frequent problem
in software development, developers already posses the tools to prevent and catch indirect conflicts to a manageable degree and have
the mindset of working until something breaks, then attempting a fix. Thus, developers would rather spend their time developing instead
of taking or reading preventative measures, resulting in the need of tools to debug and solve indirect conflicts quickly as opposed to tools 
which attempt to prevent or catch indirect conflicts early.

We continue this paper by first presenting our summarized methodology we used to perform this study in Section~\ref{sec:meth}. We
then explain and discuss the results of our study in regards to our 3 research questions in Section~\ref{sec:results}.
Next, an evaluation of how our study was performed in terms of the credibility
and quality of our results is described in Section~\ref{sec:eval}. We then cover related work in Section~\ref{sec:related}
and conclude in Section~\ref{sec:conc}.

@@

A second implication for research derived by the paper, which is more directly related to awareness, is the continued
understanding of the disjoint between theoretical awareness and awareness as perceived by industry practitioners.
As was identified in the previous discussion, while theory suggests that awareness of all events related to ones work
produces a more coherent understanding of a person's environment, developers find this to be overly time consuming
to the point where they only want to be aware of events which require action on their part. This disjoint, of what 
theoretically should be and what physically is, should be further understood to combat future potential failures in
tool development or theories attempted to be used in practice. These disjoints may be detected by conducting case 
studies of tools built primarily on theoretical models, while having these tools tested in industry. We found developers
to be quite honest when using a tool they dislike which should lend itself to the detection of the aforementioned
disjoints.

>>>>

As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
in practice we see these mitigated losses continue to appear quite frequently and can prove to be a significant
and time-consuming chore for developers~\cite{Perry:2001:PCL}.

Two types of conflicts have attracted the attention of academics, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact, or finding expert knowledge about a particular file. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. Indirect conflicts involve source code change influences such as making a change which
causes an error in another location in the system, having code negatively affected by a library upgrade, or even
breaking an integration test of a separate component than the one changed. Indirect conflict tools however, have
not shared the same success as direct conflict 
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.

While indirect conflict tools have shown potential from developer studies, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload. Through case studies, developers have noted that current indirect conflict tools provide too many 
false positive results leading to information overload and the tool eventually being
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method 
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the 
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
also compounded by the scale of many software development projects leading to further information overload.

Human factors have also shown to cause issues as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence has been leveraged in several indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
Factors again of information overload, human dependencies, and scalability lay at the root of these problems.

%Difficulties regarding indirect conflicts are to determine when software dependencies change and more crucially,
%when those changes will create negative consequences in a software project.
%Sarma et al.~\cite{Sarma:2007:TSA} built Palantir, which can both detect potential indirect conflicts, at the class signature level,
%and alert developers to these conflicts. Holmes et al.~\cite{Holmes:2010:CAR} take it
%one step further with their tool YooHoo by detecting fine grained source code changes, such as method return type changes, 
%and create a taxonomy for different
%types of changes and their proneness to cause indirect conflicts. The tool Ariadne~\cite{Trainer:2005:BGT} 
%creates an environment where developers can see how source code changes will affect other areas of a project at the 
%method level and thus where indirect conflicts may occur. 
%Another indirect conflict tool, CASI~\cite{Servant:2010:CPI},
%utilizes dependency slicing~\cite{Bajracharya:2009:SIS} instead of method call graphs to provide 
%an environment to see what areas of a project are being affected by a source code change. 
%Most of these tools have all shown to have common difficulties of: scalability, false positives, information overload, 
%or providing redundant information to preexisting tools. 
%Our own tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered these same fates.

%Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
%times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
%developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
%However, socio-technical congruence is largely unproven in regards to its correlation to
%software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.

In the hopes of resolving some of the aforementioned issues surrounding indirect conflicts and their tools, we 
are interested in exploring the following three research questions:

@@

We interviewed 19 developers from across 12 companies and open source projects as well as surveyed 78 
developers in order to answer the aforementioned questions. Our findings indicate that: indirect conflicts are very much a real 
and frequent problem in software development, contextual patterns of indirect conflicts have been missed in previous tools through
the lack of dependency identification, that developers tend to prefer to use detection and resolution processes or tools
over those of prevention, and that developers do not want awareness mechanisms which provide non actionable results.

We continue this paper by first presenting our methodology used to perform this study in Section~\ref{sec:meth}. We
then present and discuss the results of our study in regards to our 3 research questions in Section~\ref{sec:results}
and Section~\ref{sec:disc}.
Next, an evaluation of how our study was performed in terms of the credibility
and quality of our results is described in Section~\ref{sec:eval}. We then cover related work in Section~\ref{sec:related}
and conclude in Section~\ref{sec:conc}.

@@

A second implication for research derived by the paper, which is more directly related to awareness, is the continued
understanding of the disjoint between theoretical awareness and awareness as perceived by industry practitioners.
As was identified in the previous discussion, while theory suggests that awareness of all events related to ones work
produces a more coherent understanding of a person's environment, developers find this to be overly time consuming
to the point where they only want to be aware of events which require action on their part. This disjoint, of what 
theoretically should be and what physically is, should be further understood to combat future potential failures in
tool development or theories attempted to be used in practice.

These disjoints may be detected by conducting case 
studies of tools built primarily on theoretical models, while having said tools tested in industry. We found developers
to be quite honest when using a tool they dislike which should lend itself to the detection of the aforementioned
disjoints.
